Deterministic Scheduling of Periodic Messages for Cloud RAN

### 2017-07-26 infocom/infocom.tex
améliorations marginales

\begin{abstract}
A recent trend in mobile networks is to centralize in distant data-centers the processing units which were attached to 
antennas. The main challenge is to guarantee that the latency of the periodic messages sent from the antennas to their processing
units and back fulfills protocol time constraints. We show that traditional statistical multiplexing does not allow such a low latency, due to collisions and buffering at nodes. Hence, we propose in this article to use a deterministic scheme for sending periodic messages without collisions in the network thus without buffering and the resulting additional latency. 
We give several algorithms to compute such schemes for a common topology where one link is shared by all antennas.
We show that there is always a solution when the routes are short or the load is small. When the parameters are unconstrained,
and some buffering is allowed in the processing units, we propose an algorithm (PMLS) adapted from a classical scheduling method.
The experimental results show that even under full load, most of the time PMLS finds a deterministic sending scheme with no latency.
\end{abstract}

@@

Next generations of mobile network architectures evolve toward centralized radio network architectures called C-RAN, or Cloud Radio Access Network, to reduce energy consumption costs~\cite{mobile2011c} and more generally the total cost of ownership. The main challenge for this type of architecture is to reach a latency compatible with transported protocols. The latency is measured between the sending of a message by the Remote Radio Heads (RRHs) on the field and the receptions of the answers, computed by real-time virtualized network functions of BaseBand Units (BBUs), in the cloud. For example, LTE standards requires meeting time constraints for functions like HARQ (Hybrid Automatic Repeat reQuest) that needs to be processed in $3$ms~\cite{bouguen2012lte}. In 5G, some services need end to end latency as low as 1ms~\cite{3gpp5g}. The specificity of the C-RAN context is not only the latency constraint, but also the periodicity of the data transfer in the \emph{frontaul} network between RRHs and BBUs: frames need to be emitted and received every milliseconds~\cite{bouguen2012lte}.
Our aim is to operate a C-RAN on a low-cost shared switched network.
The question we address is the following : is it possible to schedule messages emission such that there is no collision, then no buffering nor latency corresponding to queuing delays ?

Let us expose briefly our model: the network topology is modeled by a directed weighted graph and a set of paths (routes) from source nodes (RRHs) to target nodes (BBUs). Time is discretized and a unit of time or slot corresponds to the transmission of a minimal unit of data over the network. The process must be  predictable (i.e. the date of arrival of packets is known beforehand).Following LTE standard~\cite{bouguen2012lte}, in this article we assume that the process is periodic. That is, if we look at the network at times $t$ and $t+P$ where $P$ is the period, all messages are at the same position in the network. As we want to avoid any  buffering in internal nodes of the graph. Hence we need to design a periodic process with no collision between messages. To ensure that this property is always true we propose a deterministic process.  We have two sets of values that we can choose when building a periodic sending process, called a \emph{periodic assignment} in this article: the time at which each packet is sent by a RRH in each period and the waiting time in the BBU before the answer is sent back to the RRH. When building a periodic assignment, we must take into account the periodicity which makes many scheduling methods inapplicable. Not only a message must not collide with the other messages sent by the others BBU/RRH in the same period, but also in the previous and following periods.The latency, that is the time between the emission of a message and the complete return of its answer, must be minimized: this lefts budget for other latency causes, and leads to larger fiber lengths, then wider deployment areas.This means that the only buffering we are allowed -- the waiting time before sending back the answer-- must be small, in particular when the route is long. Note that the
model is technology agnostic, i.e. it is compatible with an optical network with a fixed packet size.

\footnote{Others terminologies and functional splits exist in the literature. We choose the most classical one for sake of clarity. The results of this work is fully compatible with any C-RAN architecture definition.}

Our main contributions are the following.
 In Sec.~\ref{sec:def} we propose a model of the network and the periodic sending of messages along its routes  and we introduce the star routed network that we study in this article. 
 We formalize the problem of finding a periodic assignment for sending messages without buffering (\pazl) or with buffering in the processing unit only (\pall) and shows that they are $\NP$-hard even to approximate. In Sec.~\ref{sec:PAZL}, we propose two algorithms solving \pazl when the routes between RRHs and BBUs are small or when the load is light. Using a compact representation of an assignment, we show that \pazl is fixed parameter tractable when parametrized by the number of antenna on star routed networks. Finally in Sec.~\ref{sec:PALL}, we introduce a family of algorithms to solve \pall and our experimentation allows to select one which works extremely well even in very loaded networks. In particular we show that the deterministic communication schemes we design vastly outperform the traditional statistical multiplexing with regard to latency.

@@

 Statistical multiplexing even with a large bandwidth does not comply with the latency requirements of C-RAN. Therefore, the current solution~\cite{pizzinat2015things,tayq2017real} is to use dedicated circuits for the fronthaul. Each end-point (RRH on one side, BBU on the other side) is connected through direct fiber or full optical switches. This architecture is very expensive and hardly scales in the case of a mobile network composed of tenths of thousands of deployment sites. The deterministic approach we propose has gained some traction recently: Deterministic Networking is under standardization in IEEE 802.1 TSN group~\cite{finn-detnet-architecture-08}, as well at IETF DetNet working group~\cite{ieee802}. Several patents on concepts and mechanisms for DetNet have been already published, see for example~\cite{howe2005time,leclerc2016transmission}.

The algorithmic problem we focus on may look like wormhole problems~\cite{cole1996benefit}, but  we want to minimize the time lost in buffers and not just to avoid deadlocks. Several graph colorings have been introduced to model similar problems such as the allocation of frequencies~\cite{borndorfer1998frequency}, bandwidths~\cite{erlebach2001complexity} or routes~\cite{cole1996benefit} in a network or train schedules~\cite{strotmann2007railway}. Unfortunately, they do not take into account the periodicity of the scheduling and the associated problems are already $\NP$-complete. The only coloring with periodicity is the circular coloring~\cite{zhou2013multiple} but it is not expressive enough to capture our problem.

@@

\noindent \emph{Sketch of proof:}
 First, by a symmetry argument, the problem of finding an assignment with no collisions for the forward routes only
 can be reduced to \pall.
 We fix $\tau$ the size of messages to $1$ and we reduce the problem of $k$-coloring to the 
 problem of finding an assignment without collisions for the forward routes.
 Let $H$ be a graph instance of the $k$-coloring problem, from it we build a routed network $(G,{\cal R})$. 
 For each vertex $v$ in $H$, there is a route $r_v$ in ${\cal R}$. Two routes $r_v$ and $r_u$ share an arc if and only if $(u,v)$ is an arc in $H$; this arc is the only one shared by these two routes. All arcs are of weight $0$. 
 Observe that the existence of a $k$-coloring of $H$ is equivalent to the existence of a $(k,1)$-periodic assignment in $G$, 
 by converting an offset of a route into a color of a vertex and reciprocally. \qed

       We also consider a simpler version of \pall, that we call {\bf Periodic Assignment for Zero Latency} or \pazl: we ask for a $(P,\tau)$-periodic assignment {\bf with all waiting times equal to $0$}. We introduce \pazl because it is simpler to study and we are able to prove theoretical results and find better algorithms than for \pall. As we  experimentally show in Sec.~\ref{sec:exp_PAZL}, this problem can often be solved positively albeit less often than the general problem. Finally, a solution to \pazl is simpler to implement in real telecommunication networks, since we do not need any buffering at all.

@@

      \begin{figure}[h]
      \begin{center}
	 \includegraphics[width=0.47\textwidth]{periode_petite.pdf}
      \end{center}
      \caption{Minimum period averaged over $1,000$ random instances}\label{fig:short}
      \end{figure}
      First, we remark that the exhaustive search finds a solution even when the load is high, especially when there are more routes.
      It justifies the idea to look for an assignment without waiting time, in this short routes regime.
      Second, remark that the Shortest-Longest algorithm is as good as the exhaustive search. While it was expected to be good with short routes, it turns out to be optimal for all the the random star routed networks we have tried. Therefore, we should use it in practical applications with short routes, instead of the exhaustive search which is much more computationally expensive. 
      Finally, note that the greedy algorithm works on average when the load is less than $2/3$ which is twice better than the theoretical bound. This algorithm seems to depends on the load only and not on the number of routes.

@@

      In this regime, the performances of Shortest-Longest are abysmal since it depends on the difference between the longest and the smallest route which is large here. On the other hand, the greedy algorithm has a performance not so far from the case of short routes, which is expected since it does not directly depend on the size of the route. In fact, if we do the previous experiment 
      for short routes but with long routes, we find that, on average, the greedy algorithm finds a solution when the load is less than $59\%$.

      When the load is larger than $50\%$, the exhaustive search finds more solutions than the greedy algorithms which justifies its use. However, for load larger than $80\%$ there are many instances for which there are no solutions to \pazl.
      It means that with long routes and high load, looking for an assignment without waiting time is far too restrictive. That is why we present algorithms for the general \pall problem in our next section. We will test them on $8$ long routes and a load between $100\%$ and $80\%$ for which, as shown here, there are often no assignment without waiting times.

      The computation time of the exhaustive search is in $O(n!)$ as shown int Th.~\ref{th:FPT}, hence we need to study its scalability when $n$ grows. In the following experiment, the weights of the arcs $(c_t,t_i)$ are drawn following an uniform distribution between $0$ and $20,000$ slots. We chose  $95\%$ of load.
%       , that corresponds to loads of $\simeq 70\%$ (for $1$ route) to $\simeq 98\%$ (for 21 routes).
      The following table shows the time before the exhaustive search ends, for $8$ to $16$ routes, averaged on $100$ random star routed networks. This shows that for less than $15$ routes, which corresponds to all current topologies, the algorithm is efficient enough, but we should improve it further to work on more routes.
         \begin{center}
         \begin{tabularx}{0.5\textwidth}{|l|X|X|X|X|X|}
    \hline
   $n$ & $8$ & $10$& $12$&$14$& $16$\\
    \hline
   Time (s) & $6.10^{-5}$&$8.10^{-4}$&$2.10^{-2}$& $0.4$& $11$\\
    \hline
      \end{tabularx}
      \end{center}

@@

        We can simplify the weights on the arcs $(s_i,c_s)$ by setting them all to $0$.
        We find the route with maximal length, say that it is $\lambda(r_0)$. The idea is to 
        set the waiting times of all routes so they behave exactly as the message in $r_0$.

@@

    We can decompose any algorithm solving the \pall problem in two stages: first set the offsets of the forward routes and then, knowing this information set the offset of the backward routes or equivalently the waiting times.  
    The offsets of the forward routes are chosen so that they have no collisions and such that there are no free slots in between the end of a message on the central arc and the beginning of the next one. 
    It is done to minimize the period needed to send the messages of the forward routes. For this section and the next, we assume that the offsets of the forward routes are fixed by some order.

@@

    The first algorithm we propose is a greedy algorithm which sets the offset $m_{\rho(r_i)}$ of the 
    backward routes. It prioritizes the routes with the earliest deadline to best satisfy the
    constraint on the process time. Set $t=0$ and repeat the following: find $s \geq t$ the first time for which there is an eligible route with its offset not fixed. Then amongst all eligible routes at time $s$ choose the one with the smallest deadline, fix its offset to $s - \omega(c_t,t_i) $ and set $t = s + \tau$.

@@

    The greedy algorithm can be made to work in time $O(n\log(n))$ by using a binary heap for the set of eligible routes, and a list to maintain a list of the intervals
    of time at which a message can be sent on the arc $(c_t,c_s)$.
%    To do that the set of eligible routes must be maintained in a binary heap
%    to be able to find the one with smallest deadline in time $O(\log(n))$. 
%    To deal with the possible collisions, one can maintain a list of the intervals
%    of time at which a message can be sent on the arc $(c_t,c_s)$. Each time the offset of a 
%    route is fixed an interval is split into at most two intervals in constant time. 
%    Since the algorithm goes over the elements of the list at most twice when doing an insertion
%    or looking for the next free interval, the time needed to maintain it is $O(n)$. 
%    
%     
%          We compared the two algorithms on $10000$ graphs on which the weight of the links are drawn uniformly between $0$ and $20000$, with a period of $21000$, and giving, for each instance $T_{max} = 2\lambda(r_{longest-route})$ . We then look at the success rate of each algorithms, with every sending order. Both algorithms reach the higher success rate with the Random sending order, but the success of Greedy is $5,65 \%$, while the success rate of Greedy Periodic is to $30,02 \%$.

@@

     The problem to solve in the second stage of our approach is very similar to the following scheduling problem: 
     we are given a set of jobs and each job has a \emph{release time} and a \emph{deadline}. 
     The problem is to schedule all jobs on a single processor, that is choosing the time at which they are computed, so that no two jobs are scheduled at the same time. A job is always scheduled after its release time and it must be finished before its deadline. Let us call $n$ the number of jobs, the problem can be solved in time $O(n^2\log(n))$~\cite{simons1978fast} when all jobs have the same running time and it gives a solution with the earliest possible deadline. On the other hand if the running times are different the problem is $\NP$-complete~\cite{lenstra1977complexity}. 
     The  polynomial time algorithm  which solves this scheduling problem is similar to the Greedy algorithm presented in the previous section. However, when it fails because a job finishes after its deadline, it changes the schedule of the last messages to find a possible schedule for the problematic job. The change in the scheduling is so that the algorithm cannot fail on the same job a second time except if there is no solution, which proves that the algorithm is in polynomial time.

@@

    We fix the number of routes to $8$ to make comparisons with the results of Section~\ref{sec:exp_PAZL} easier. 
    We draw uniformly the weights of the arcs between $0$ and $20,000$.
    For a given routed network, we define the {\em margin} as the difference between $T_{max}$ and twice the longest route. 
    Hence, for a given routed network choosing the margin or $T_{max}$ is the same, but since we will consider many networks whith different longest route's size, it will be more meaningful to set a global margin rather than $T_{max}$.
    In fact the margin represents the latency imposed by the communication process without taking into account the physical length of the network which cannot be changed. In our experiments the margin range from  $0$ to $3,000$.
   We look at two different regimes, a medium load of $80\%$ and a high load of $95\%$.
   Considering smaller load is not relevant since then we know how to solve the problem even without waiting times as shown in Section~\ref{sec:exp_PAZL}.

@@

    \begin{figure}
       \begin{center}
      \includegraphics[width = 0.45\textwidth]{stochastic.pdf}
       %\input{figures/stochastic}
      \end{center}
      \vspace{-1cm}
      \caption{Cumulative distribution of the margin of statistical multiplexing}
      \label{fig:sto}
     \end{figure}

     We have also computed the minimum margin needed by PMLS to find a solution but we have not represented it in Figure~\ref{fig:sto} since even the third quartile was always at $0$ for a load of $100\%$. There is around $20\%$ of the instances for which the margin is larger than $0$ for a load of $100\%$ and $2\%$ for a load of $95\%$. For the instances with a margin larger than $0$, the margin is on average $400$.

>>>>

\begin{abstract}
A recent trend in mobile networks is to centralize in distant data-centers the processing units which were attached to antennas. The main challenge is to guarantee that the latency of the periodic messages sent from the antennas to their processing units and back fulfills protocol time constraints. We show that statistical multiplexing does not allow such a low latency, due to collisions and buffering at nodes. Hence, we propose to use a deterministic scheme for sending periodic messages without collisions in the network to avoid buffering and the resulting additional latency. We give several algorithms to compute such schemes for a common topology where one link is shared by all antennas. We show that there is always a solution when the routes are short or the load is small. When the parameters are unconstrained, and some buffering is allowed in the processing units, we propose an algorithm (PMLS) adapted from a classical scheduling method. The experimental results show that even under full load, PMLS finds most of the time a deterministic sending scheme with no latency.
\end{abstract}

@@

Next generations of mobile network architectures evolve toward centralized radio network architectures called C-RAN for Cloud Radio Access Network, to reduce energy consumption costs~\cite{mobile2011c} and more generally the total cost of ownership. The main challenge for this type of architecture is to reach a latency compatible with transport protocols. The latency is measured between the sending of a message by a Remote Radio Head (RRH) and the receptions of the answer, computed by real-time virtualized network functions of a BaseBand Unit (BBU) in the cloud. For example, LTE standards requires to process functions like HARQ (Hybrid Automatic Repeat reQuest) in $3$ms~\cite{bouguen2012lte}. In 5G, some services need end to end latency as low as $1$ms~\cite{3gpp5g}. The specificity of the C-RAN context is not only the latency constraint, but also the periodicity of the data transfer in the \emph{frontaul} network between RRHs and BBUs: frames need to be emitted and received each millisecond~\cite{bouguen2012lte}.
Our aim is to operate a C-RAN on a low-cost shared switched network.
The question we address is the following: is it possible to schedule messages such that there is no collisions to avoid latency caused by queuing delays? 
Eliminating this source of latency leaves us with more time budget for latency coming from the physical length of the routes in the network, and thus allows for wider deployment areas.

Let us expose briefly our model: the network topology is modeled by a directed weighted graph and a set of paths (routes) from source nodes (RRHs) to target nodes (BBUs). Time is discretized and a unit of time or slot corresponds to the transmission of a minimal unit of data over the network. The process must be  predictable (i.e. the date of arrival of packets is known beforehand). Following LTE standard~\cite{bouguen2012lte}, in this article we assume that the process is periodic. That is, if we look at the network at times $t$ and $t+P$ where $P$ is the period, all messages are at the same position in the network. As we want to avoid any  buffering in internal nodes of the graph. Hence we need to design a periodic process with no collision between messages. To ensure that this property is always true we propose a deterministic process.  We have two sets of values that we can set when building a periodic sending process, called a \emph{periodic assignment} in this article: the time at which each packet is sent by a RRH in each period and the waiting time in the BBU before the answer is sent back to the RRH. When building a periodic assignment, we must take into account the periodicity which makes many scheduling methods unusable. Not only a message must not collide with the other messages sent by the others BBU/RRH in the same period, but also in the previous and following periods. The latency, that is the time between the emission of a message and the complete return of its answer must be minimized. This means that the only buffering we are allowed -- the waiting time before sending back the answer-- must be small, in particular when the route is long. Note that the model is technology agnostic, i.e. it is compatible with an optical network with a fixed packet size.

%\footnote{Others terminologies and functional splits exist in the literature. We choose the most classical one for sake of clarity. The results of this work is fully compatible with any C-RAN architecture definition.}

Our main contributions are the following.
 In Sec.~\ref{sec:def} we propose a model of the network and the periodic sending of messages along its routes  and we introduce the star routed network that we study in this article. 
 We formalize the problem of finding a periodic assignment for sending messages without buffering (\pazl) or with buffering in the processing unit only (\pall) and shows that they are $\NP$-hard even to approximate. In Sec.~\ref{sec:PAZL}, we propose two algorithms solving \pazl when the routes between RRHs and BBUs are small or when the load is light. Using a compact representation of an assignment, we show that \pazl is fixed parameter tractable when parametrized by the number of antennas on star routed networks. Finally in Sec.~\ref{sec:PALL}, we introduce a family of algorithms to solve \pall and our experimentations allow to select one which works extremely well even in loaded networks. In particular we show that the deterministic communication schemes we design vastly outperform the traditional statistical multiplexing with regard to latency.

@@

 Statistical multiplexing even with a large bandwidth does not comply with the latency requirements of C-RAN. Therefore, the current solution~\cite{pizzinat2015things,tayq2017real} is to use dedicated circuits for the fronthaul. Each end-point, RRH or BBU, is connected through direct fiber or full optical switches. This architecture is very expensive and hardly scales in the case of a mobile network composed of tenths of thousands of deployment sites. The deterministic approach we propose has gained some traction recently: Deterministic Networking is under standardization in IEEE 802.1 TSN group~\cite{finn-detnet-architecture-08}, and in  IETF DetNet working group~\cite{ieee802}. Several patents on concepts and mechanisms for DetNet have been already published, see for example~\cite{howe2005time,leclerc2016transmission}.

The algorithmic problem we focus on may look like wormhole problems~\cite{cole1996benefit}, but  we want to minimize the time lost in buffers and not just to avoid deadlocks. Several graph colorings have been introduced to model similar problems such as the allocation of frequencies~\cite{borndorfer1998frequency}
% bandwidths~\cite{erlebach2001complexity}
or routes~\cite{cole1996benefit} in a network or train schedules~\cite{strotmann2007railway}. Unfortunately, they do not take into account the periodicity of the scheduling and the associated problems are already $\NP$-complete. The only coloring with periodicity is the circular coloring~\cite{zhou2013multiple} but it is not expressive enough to capture our problem.

@@

\noindent \emph{Sketch of proof:}
 First, by a symmetry argument, the problem of finding an assignment with no collisions for the forward routes only
 can be reduced to \pall.
 We fix $\tau$ the size of messages to $1$ and we reduce the problem of $k$-coloring to the 
 problem of finding an assignment without collisions for the forward routes.
 Let $H$ be a graph instance of the $k$-coloring problem, from it we build a routed network $(G,{\cal R})$. 
 For each vertex $v$ in $H$, there is a route $r_v$ in ${\cal R}$. Two routes $r_v$ and $r_u$ share an arc if and only if $(u,v)$ is an arc in $H$; this arc is the only one shared by these two routes. All arcs are of weight $0$. 
 Observe that the existence of a $k$-coloring of $H$ is equivalent to the existence of a $(k,1)$-periodic assignment in $G$, 
 by converting an offset of a route into a color of a vertex.\qed
%  and reciprocally. \qed

       We also consider a simpler version of \pall, that we call {\bf Periodic Assignment for Zero Latency} or \pazl: we ask for a $(P,\tau)$-periodic assignment {\bf with all waiting times equal to $0$}. We introduce \pazl because it is simpler to study and we are able to prove theoretical results and find better algorithms than for \pall. As we  experimentally show in Sec.~\ref{sec:exp_PAZL}, this problem can often be solved positively albeit less often than the general problem. Finally, a solution to \pazl is simpler to implement in real telecommunication networks, since we do not need buffering at all.

@@

      \begin{figure}[h]
      \begin{center}
	 \includegraphics[width=0.47\textwidth]{periode_petite.pdf}
      \end{center}
      \caption{Minimum period averaged over $1,000$ random instances}\label{fig:short}
      \end{figure}
      First, we remark that the exhaustive search finds a solution even when the load is high, especially when there are more routes.
      It justifies the idea to look for an assignment without waiting time, in this short routes regime.
      Second, remark that the Shortest-Longest algorithm is as good as the exhaustive search. While it was expected to be good with short routes, it turns out to be optimal for all the the random star routed networks we have tried. Therefore, we should use it in practical applications with short routes, instead of the exhaustive search which is much more computationally expensive. 
      Finally, note that the greedy algorithm works on average when the load is less than $2/3$ which is twice better than the theoretical bound. This algorithm seems to depend on the load only and not on the number of routes.

@@

      In this regime, the performances of Shortest-Longest are abysmal since it depends on the difference between the longest and the smallest route which is large here. On the other hand, the greedy algorithm has a performance not so far from the case of short routes, which is expected since it does not directly depend on the size of the route. In fact, if we do the experiment of Fig.~\ref{fig:short} for long routes, on average the greedy algorithm finds a solution when the load is less than $59\%$.

      When the load is larger than $50\%$, the exhaustive search finds more solutions than the greedy algorithms which justifies its use. However, for loads larger than $80\%$ there are many instances for which there are no solutions to \pazl: \emph{with long routes and high load}, looking for an assignment without waiting time is far too restrictive. That is why we present algorithms for the general \pall problem in the next section. We will test them on $8$ long routes and a load between $100\%$ and $80\%$ for which, as shown here, there are often no assignments without waiting times.

      The computation time of the exhaustive search is in $O(n!)$ as shown int Th.~\ref{th:FPT}, hence we need to study its scalability when $n$ grows. We set the load at $95\%$.
%       
%       In the following experiment, the weights of the arcs $(c_t,t_i)$ are drawn following an uniform distribution between $0$ and $20,000$ slots. We chose  $95\%$ of load.
%       , that corresponds to loads of $\simeq 70\%$ (for $1$ route) to $\simeq 98\%$ (for 21 routes).
      The following table shows the time before the exhaustive search ends, for $8$ to $16$ routes, averaged on $100$ random star routed networks drawn as in the previous experiment. It shows that for less than $15$ routes, which corresponds to all current topologies, the algorithm is efficient enough, but it should be further improved to work on more routes.
         \begin{center}
         \begin{tabularx}{0.5\textwidth}{|l|X|X|X|X|X|}
    \hline
   $n$ & $8$ & $10$& $12$&$14$& $16$\\
    \hline
   Time (s) & $6.10^{-5}$&$8.10^{-4}$&$2.10^{-2}$& $0.4$& $11$\\
    \hline
      \end{tabularx}
      \end{center}

@@

        We can simplify the weights on the arcs $(s_i,c_s)$ by setting them all to $0$.
        Let assume that the route with maximal length is $r_0$. The idea is to 
        set the waiting times of all routes so they behave exactly as the message in $r_0$.

@@

    We can decompose any algorithm solving the \pall problem in two stages: first set the offsets of the forward routes and then set the offset of the backward routes or equivalently the waiting times.  
    The offsets of the forward routes are chosen so that they have no collisions and such that there are no free slots in between the end of a message on the central arc and the beginning of the next one. It is done to minimize the period needed to send the messages of the forward routes. Remark that an order on the routes is enough to determine the offsets of the forward routes with the previous property. In this subsection and the next, we assume that the offsets of the forward routes are fixed by some order.

@@

    The first algorithm we propose is a greedy algorithm which sets the offset $m_{\rho(r_i)}$ of the backward routes. It prioritizes the routes with the earliest deadline to best satisfy the
    constraint on the process time, hence we call it \textbf{Greedy Deadline (GD)}. Set $t=0$ and repeat the following: find $s \geq t$ the first time for which there is an eligible route with its offset not fixed. Then amongst all eligible routes at time $s$ choose the one with the smallest deadline, fix its offset to $s - \omega(c_t,t_i) $ and set $t = s + \tau$.

@@

    Algorithm GD can be made to work in time $O(n\log(n))$ by using a binary heap for the set of eligible routes, and a list to maintain a list of the intervals of time at which a message can be sent on the arc $(c_t,c_s)$.
%    To do that the set of eligible routes must be maintained in a binary heap
%    to be able to find the one with smallest deadline in time $O(\log(n))$. 
%    To deal with the possible collisions, one can maintain a list of the intervals
%    of time at which a message can be sent on the arc $(c_t,c_s)$. Each time the offset of a 
%    route is fixed an interval is split into at most two intervals in constant time. 
%    Since the algorithm goes over the elements of the list at most twice when doing an insertion
%    or looking for the next free interval, the time needed to maintain it is $O(n)$. 
%    
%     
%          We compared the two algorithms on $10000$ graphs on which the weight of the links are drawn uniformly between $0$ and $20000$, with a period of $21000$, and giving, for each instance $T_{max} = 2\lambda(r_{longest-route})$ . We then look at the success rate of each algorithms, with every sending order. Both algorithms reach the higher success rate with the Random sending order, but the success of Greedy is $5,65 \%$, while the success rate of Greedy Periodic is to $30,02 \%$.

@@

     The problem to solve in the second stage of our approach is very similar to the following scheduling problem: 
     we are given a set of jobs and each job has a \emph{release time} and a \emph{deadline}. 
     The problem is to schedule all jobs on a single processor, that is choosing the time at which they are computed, so that no two jobs are scheduled at the same time. A job is always scheduled after its release time and it must be finished before its deadline. Let us call $n$ the number of jobs, the problem can be solved in time $O(n^2\log(n))$~\cite{simons1978fast} when all jobs have the same running time and it gives a solution with the earliest possible deadline. On the other hand if the running times are different the problem is $\NP$-complete~\cite{lenstra1977complexity}. 
     The  polynomial time algorithm  which solves this scheduling problem is similar to the greedy algorithm presented in the previous section. However, when it fails because a job finishes after its deadline, it changes the schedule of the last messages to find a possible schedule for the problematic job. The change in the scheduling is so that the algorithm cannot fail on the same job a second time except if there is no solution, which proves that the algorithm is in polynomial time.

@@

    We fix the number of routes to $8$ to make comparisons with the results of Section~\ref{sec:exp_PAZL} easier. 
    We draw uniformly the weights of the arcs between $0$ and $20,000$.
    For a given routed network, we define the {\em margin} as the difference between $T_{max}$ and twice the longest route. 
    Hence, for a given routed network choosing the margin or $T_{max}$ is the same, but since we will consider many networks with different longest route's size, it will be more meaningful to set a global margin rather than $T_{max}$.
    In fact the margin represents the latency imposed by the communication process without taking into account the physical length of the network which cannot be changed. In our experiments the margin range from  $0$ to $3,000$.
   We look at two different regimes, a medium load of $80\%$ and a high load of $95\%$.
   Considering smaller load is not relevant since then we know how to solve the problem even without waiting times as shown in Section~\ref{sec:exp_PAZL}.

@@

    \begin{figure}
       \begin{center}
      \includegraphics[width = 0.45\textwidth]{stochastic.pdf}
       %\input{figures/stochastic}
      \end{center}
      \caption{Cumulative distribution of the margin of statistical multiplexing}
      \label{fig:sto}
     \end{figure}

     We have also computed the minimum margin needed by PMLS to find a solution but we have not represented it in Fig.~\ref{fig:sto} since even the third quartile was always at $0$ for a load of $100\%$. There is around $20\%$ of the instances for which the margin is larger than $0$ for a load of $100\%$ and $2\%$ for a load of $95\%$. For the instances with a margin larger than $0$, the margin is on average $400$.
